---
title: "Examen Topicos Selectos de Software"
author: "Daniela Torres Lopez"
output: html_notebook
---

##1.- Create a numeric vector that contains the eight planets of our solar system. Then, create another vector containing the name for each planet (Mercury, Venus, Earth, Mars, Jupiter, Saturn, Uranus, and Neptune). After that, you should assign the number of the planet and its respective name. Finally, you should select and print the planets related to the names Earth, Mars, and Jupiter.

```{r}
planet_numbers <- c(1:8)
planet_names <-c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
names(planet_numbers) <- planet_names
print(planet_numbers[3:5])
```


##2.- Construct a matrix with 10 rows containing the numbers 1 up to 50, filled row-wise.
```{r}
mat <- matrix(c(1:50), nrow = 10, byrow = TRUE)
print(mat)
```


##3.-Create a factor with these observations "Breakfast", "Breakfast", "Breakfast", "Lunch", "Lunch", "Dinner", "Dinner", "Dinner", "Collation", "Collation", "Collation" and print a summary.
```{r}

obs <- c("Breakfast", "Breakfast", "Breakfast", "Lunch", "Lunch", "Dinner", "Dinner", "Dinner", "Collation", "Collation", "Collation")

obs_factor <- factor(obs)
summary(obs_factor)

```


##4.- Create and print a data frame that should contain:

5 of your courses (e.g. Fundamentals of Programming, Object-Oriented Programming, Databases, etc ).
5 of your marks
5 of your professors' names

```{r}
courses <- c("Arquitectura de Software", "Topicos Selectos", "Innovacion Tecnologica", "IHM", "Arquitectura de las Computadoras")

Marks <- c(8, 10, 7,7,7)

Professors <- c("Vicente Garcia", "Alan Ponce", "Abraham Lopez", "Francisco Lopez", "Oscar Martinez")

dataClasses <- data.frame(courses, Marks, Professors)
dataClasses
```


##5.- Load the dataset labeled as odb.csv that is located in the data folder and storage this dataset in a variable called od. Then, you should describe the composition of the dataset. This description must contain the number of observations, variables and the type of each variable.
35 Variables
549 Obs.
Type:
  V1: Num
  V2: Char
  V3: Num
  V4: Char
  V5: Char
  v6: Char
  V7: Char
  V8: Char
  V9:V35: Num Dbl

```{r}
 od<-odb
summary(od)
lengths(od)

```



##6.- Filter the od dataset to retrieve only the observation from Mexico, grouped by year and in descending order selecting the "ODB_Rank column
```{r}
filtered<-od
  filter(Economy=="Mexico") 
  group_by(year)
  arrange(desc(ODB_Rank))
  print(filtered)
```


##7.- Create a new variable called od_year grouping per "year" and "Income group" variables. Then, remove NAs observations. Later, using the function summarize() you should estimate the median of the GDP_PPP variable and store it in a variable called "medianValue"
```{r}
od_year <- od%>% group_by(year,Income.group)%>% filter() %>% summarise(medianValue = median(GDP_PPP)) 
  
od_year <- od %>% group_by(year,Income.group) 
  print(od_year) 

```


##8.-Create a line plot graph to visualize trends over time. You should use the variables created in the od_year data frame using Income group as color
```{r}
ggplot(itu_by_year, aes(x=year, y=medianValue, color=Income.group)) + geom_line() + ggtitle("Median value over time")
```


##9.- Create a new variable called od_latin filtering per Region, selecting Latin America & Caribbean and grouping per Income group. Then, using this data frame, create a plot visualizing the relationship between Individuals_using_Internet and GDP_PPP colored per Income group and faceting per Economy
```{r}
  od_latin <- od %>% filter(Region == "Latin America & Caribbean") %>% group_by(Income.group)
  print(od_latin)
  od_latin <- remove_missing(itu_latin, na.rm = TRUE);
...
```


##10.- Create a new variable called od_2016 filtering the year 2016 and removing observations that contain NAs. Then, using this data frame, create a plot visualizing the relationship between Technological adoption and GDP_PPP, colored per Income group. Furthermore, you should use a log scale in both axes ("x" and "y"). Finally, you should use the facet function in order to wrap by Region.
```{r}
od_2016 <- od %>% filter(year==2016)
od_2016 <- remove_missing(od_2016, na.rm = TRUE);
ggplot(od_2016, aes(x=ODB_Rank, y=GDP..US..,color=Income.group)) + geom_point() + scale_x_log10() + scale_y_log10() + facet_wrap(~Region)
```

